---
title: "Assignment 3"
author: "Ryan McGuinness"
date: "10/15/2021"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Problem description
We need to minimize total cost of producing AEDs in two plants and of shipping from those plants to three warehouses. The table summarizing the costs is shown below.

| Unit Shipping Costs | Warehouse 1 | Warehouse 2 | Warehouse 3 |   | Unit Production Cost | Monthly Capacity |
|---------------------|-------------|-------------|-------------|---|----------------------|------------------|
| Plant A             | \$22        | \$14        | \$30        |   | \$600                | 100              |
| Plant B             | \$16        | \$20        | \$24        |   | \$625                | 120              |
|                     |             |             |             |   |                      |                  |
| Monthly Demand      | 80          | 60          | 70          |   |                      |                  |

First, we notice that total supply is not equal to total demand. A dummy warehouse with a demand of 10 units monthly is needed to correct this to equality.

| Unit Shipping Costs | Warehouse 1 | Warehouse 2 | Warehouse 3 | Warehouse 4(D) |   | Unit Production Cost | Monthly Capacity |
|---------------------|-------------|-------------|-------------|----------------|---|----------------------|------------------|
| Plant A             | \$22        | \$14        | \$30        | \$0            |   | \$600                | 100              |
| Plant B             | \$16        | \$20        | \$24        | \$0            |   | \$625                | 120              |
|                     |             |             |             |                |   |                      |                  |
| Monthly Demand      | 80          | 60          | 70          | 10             |   |                      |                  |

This can be represented by the following linear program:  
$\textbf{MIN } Z = 622x_{A1} + 614x_{A2} + 630x_{A3} + 641x_{B1} + 645x_{B2} + 649x_{B3}$ subject to  
$x_{A1} + x_{A2} + x_{A3} = 100$,  
$x_{B1} + x_{B2} + x_{B3} = 120$,  
$x_{A1} + x_{B1} = 80$,  
$x_{A2} + x_{B2} = 60$,  
$x_{A3} + x_{B3} = 70$,  
$x_{A4} + x_{B4} = 10$, and  
$x_{i,j} \geq 0$.  

The formulation is included in the written up in the AED.lp file.
```{r}
library(lpSolveAPI)
AED <- read.lp("AED.lp")
AED
```

Now we solve the problem.
```{r}
solve(AED)
get.objective(AED)
get.variables(AED)
```

The last two values given are for the dummy variables $x_{A4}$ and $x_{B4}$. Our model tells us that out of the 100 units we can produce at Plant A, 60 should be set to warehouse 2 and 40 to warehouse 3, and out of the 120 units we can produce at Plant B, we only produce 110, send 80 to warehouse 1 and 30 to warehouse 3.
