---
title: "DEA"
author: "Ryan McGuinness"
date: "10/29/2021"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Problem Description
The Hope Valley Health Care Association operates six nursing homes. Each of these facilities are evaluated on the efficiency using average hours per day and cost of supplies per day as inputs and number of patient-days reimbursed by third party sources and number of patient-days reimbursed privately as outputs. A summary of performance is shown in the table below.  

| Nursing Home | Staff Hours per Day | Supply Cost per Day (thousands of dollars) | Reimbursed Patient-Days | Privately Paid Patient-Days |
|----------------------|---------------------|--------------------------------------------|-------------------------|-----------------------------|
| Facility 1           | 150                 | 0.2                                        | 14000                   | 3500                        |
| Facility 2           | 400                 | 0.7                                        | 14000                   | 21500                       |
| Facility 3           | 320                 | 1.2                                        | 42000                   | 10500                       |
| Facility 4           | 520                 | 2.0                                        | 28000                   | 42000                       |
| Facility 5           | 350                 | 1.2                                        | 19000                   | 25000                       |
| Facility 6           | 320                 | 0.7                                        | 14000                   | 15000                       |

## Preparing DEA Analysis

### Importing Required Packages
```{r}
library(Benchmarking)
```

### Reading Input Data
The `dea` function requires the input and output data to be in matrix format. So, we must start by putting the data into matrix form and give descriptive column names to those matrices.
```{r}
inputs  <- matrix(c(150, 400, 320, 520, 350, 320,
                    0.2, 0.7, 1.2, 2.0, 1.2, 0.7), ncol=2) # Defining inputs
outputs <- matrix(c(14000, 14000, 42000, 28000, 19000, 14000,
                    3500, 21000, 10500, 42000, 25000, 15000), ncol=2) # Defining outputs

# Defining column names
colnames(inputs)  <- c("Daily Staff Hours", "Daily Supplies Cost") 
colnames(outputs) <- c("Reimbursed Patient-Days", "Privately-Paid Patient Days")

inputs
outputs
```


### Creating DEA Analysis Function
To save ourselves from repetitive code, let's define a function that will perform the DEA under the given assumption.
```{r}
analyze_dea <- function(assumption) {
  analysis <- dea(inputs, outputs, RTS=assumption)
  print(eff(analysis))
  print(peers(analysis))
  print(lambda(analysis))
}
```

In order to use the function, we just supply the DEA assumption to use and the function will print the efficiencies, peers, and lambdas.

## Performing DEA Analysis
Now, we use the function created above to perform DEA on each of the six assumptions in the order of FDH, CRS, VRS, IRS, DRS, and FRH.

### Free Disposability Hull
```{r}
analyze_dea("FDH")
```

### Constant Return to Scale
```{r}
analyze_dea("CRS")
```

### Varying Return to Scale
```{r}
analyze_dea("VRS")
```

### Increasing Return to Scale
```{r}
analyze_dea("IRS")
```

### Decreasing Return to Scale
```{r}
analyze_dea("DRS")
```

### Free Replicability Hull
```{r}
analyze_dea("ADD")
```

## Summary of Results

The efficiency of each firm under the given assumption is given below.  

| DMU | $E^{FDH}$ | $E^{CRS}$   | $E^{VRS}$   | $E^{IRS}$   | $E^{DRS}$   | $E^{FRH}$ |
|----------------------|-----|-------|-------|-------|-------|-----|
| Facility 1           | 1   | 1     | 1     | 1     | 1     | 1   |
| Facility 2           | 1   | 1     | 1     | 1     | 1     | 1   |
| Facility 3           | 1   | 1     | 1     | 1     | 1     | 1   |
| Facility 4           | 1   | 1     | 1     | 1     | 1     | 1   |
| Facility 5           | 1   | 0.977 | 1     | 1     | 0.977 | 1   |
| Facility 6           | 1   | 0.867 | 0.896 | 0.896 | 0.867 | 1   |

Under all six assumptions, facilities 1, 2, 3, and 4 are considered efficient, and without the assumption of convexity all six facilities are efficient. 

Under the VRS and IRS assumptions, facility 6 is inefficient. Its peers are facilities 1, 2, and 5 with $\lambda_1^6 = 0.401$, $\lambda_2^6 = 0.342$, and $\lambda_5^6 = 0.256$.

Under the CRS and DRS assumptions, facilities 5 and 6 are both inefficient. The peers of both facilities 5 and 6 are facilities 1, 2, and 4, with $\lambda_1^5 = 0.2$, $\lambda_2^5 = 0.080$, $\lambda_3^5 = 0.538$, $\lambda_1^6 = 0.343$, $\lambda_2^6 = 0.395$, and $\lambda_3^6 = 0.131$.
